working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-02-25 Manuscript_draft/Manuscript_draft.Rmd
changes

---
title: working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  word_document: null
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---

@@

We used the statistical language `R` `r citep(citation())` for all our analyses. Differences of seed set between treatments and hand cross pollination for each species were tested through mixed linear models. For the following analysis we scaled the values of seed production for all the species with mean 0 and sd of 1 in order to be able to compare across species. To test the effect of heterospecific pollen, we substracted to the mean seed set of hand cross pollination the mean seed set of heterospecific pollen treatments. In order to see correlations between heterospecific pollen effect and traits we performed Mantel test between the matrix of heterospecific pollen effect and the distance matrix of each trait (euclidean distances). Moreover, Mantel test was also conducted between heterospecific pollen effect and the square root of the matrix of phylogenetic distance due to improvement in the statistical power (Letten & Cornwell 2014). [all is here, but I would break it by questions and give a bit more detail, to avoid overwhelm the reader]() We explored also the relations between traits and heterospecific pollen effect through generalized mixed models where the response variable was heterospecific pollen effect, the independent variable the different traits and the random effects the different treatments per species [Here I think you should think if this controls for the non independency of donors and recipients. I think not. Maybe look onto matrix regresions?). Moreover, pairwise evolutive distances were calculated with MEGA7 for two kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL). The sequences of interest were downloaded from NCBI GenBank and the phylogenetic tree constructed by maximum likelihood with MEGA7.[Make a section on how you contrsucted phylogeny]().

To evaluate heterospecific pollen effect on seed production we performed linear mixed models. The distinct heterospecific pollination treatments were compared with the cross pollination treatment which was our control for optimum seed production for all the species. The different replicates of each treatment were considered as random effects. The changes of seed production per average number of ovules across species were compared through scaling the seed production with mean 0 and standard deviation of 1. All the analysis were conducted with the statistical language `R` `r citep(citation())`

1) Mantels: relative effects

2) GLM's or matrix models

>>>>

---
title: working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  word_document: null
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---

@@

 Moreover, Mantel test was also conducted between heterospecific pollen effect and   [Here I think you should think if this controls for the non independency of donors and recipients. I think not. Maybe look onto matrix regresions?). Moreover, pairwise evolutive distances were calculated with MEGA7 for two kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL). The sequences of interest were downloaded from NCBI GenBank and the phylogenetic tree constructed by maximum likelihood with MEGA7.[Make a section on how you contrsucted phylogeny]().

To evaluate heterospecific pollen effect on seed production we performed linear mixed models. The distinct heterospecific pollination treatments were compared thorugh relevelling each variable with the cross pollination treatment which was our control for optimum seed production for all the species. The different replicates of each treatment were considered as random effects. The changes of seed production across species were compared through scaling the seed production with mean 0 and standard deviation of 1 prior to the analysis. All the analysis were conducted with the statistical language `R` `r citep(citation())`

In order to test the relative effect of traits on seed production with foreign pollen we performed mantel test between the assymetrical matrix of heterospecific pollen effect (10 by 10 matrix) with the different distance matrices of traits (euclidean distances). Heterospecific pollen effect was considered as the subtraction of seed production by hand cross pollination minus seed production of the different heterospecific pollen treatments. Moreover, to find a model with the best explanatory traits we used the function **bioenv** from R. Finally, we also conducted mantel test between the matrix of heterospecific pollen effect and the distance matrix from all the traits.

We conducted mantel test to check for correlations between heterospecific pollen effect and phylogenetic distance.

the square root of the matrix of phylogenetic distance due to improvement in the statistical power (Letten & Cornwell 2014).

??We explored also the relations between traits and heterospecific pollen effect through generalized mixed models where the response variable was heterospecific pollen effect, the independent variable the different traits and the random effects the different treatments per species

### 2019-02-25 Manuscript_draft/Manuscript_draft.tex
changes

Pollinator sharing can have negative consequences for plant fitness with
the arrival of foreign pollen. However, the costs of heterospecific
pollen are not yet well understood. We conducted a glasshouse experiment
to understand how phylogenetic relatedness and plant traits mediate the
impacts of heterospecific pollen transfer. We conducted 1800 crosses by
experimentally transferring pollen (50\% and 100\% ratio) with
reciporcal crosses between 10 species belonging to three different
families: Brassicaceae, Solanaceae and Convolvulaceae. Seed set was used
as proxy of plant fitness. We found that for 65\% of the treatments with
50\% mix reduced seed set. Moreover, the reduction in seed set was
dependent on the degree of relatedness and reproductive traits of the
pollen recipient and not the pollen donor. Our results show that certain
traits, particularly compatibility system, are critical in understanding
the costs of heterospecific pollen. \vspace{3mm}\hrule

@@

The study was conducted in a glasshouse at University of New England
(Armidale, Australia) from November 2017 to March 2018. Rooms were
temperature controlled depending on the requirements of the species with
day and night temperature differences. The experimental design had
species from three different families: Solanaceae, Brassicaceae and
Convolvulaceae (\textbf{Table 1}). The different species of the study
have different reproductive traits and different degree of relatedness
where the reciprocal crosses between species allowed us to have multiple
different scenarios of both traits and relatedness. Moreover, the
species selected have fast life cycle and low structural flower
complexity in order to perform the pollination treatments and grow the
different species from seeds. For the purpose of the experiment all the
species where considered as pollen recipient and as pollen donor (see
interaction matrix, fig 2). Species were watered once or twice per day
and fertilized weekly (NPK 23: 3.95: 14).

@@

The traits measured for each species were pollen per anther, number of
ovules, stigma width and length and stigmatic area, style width and
length, ovary width and length. Moreover stigma type \href{}{explain}
was tested. Pollen was counted for 20 anthers of each species with 4
replicates per sample with an hemocytometer. Previously, anthers were
squashed on a known solution with the pippete tip and homogeneize with a
vortex for 30 seconds. Ovule number was counted with the help of an
stereomicroscope and a small grid over a petri dish from 15 randomly
selected flowers. The different morphometrical traits were measured with
a digital stereomicrospe. Levels of self incompatibility were estimated
by dividing the fruit set of hand self pollination by hand cross
pollination Lloyd and Schoen (1992).

@@

We used the statistical language \texttt{R} (R Core Team 2018) for all
our analyses. Differences of seed set between treatments and hand cross
pollination for each species was tested through mixed linear models. For
the following analysis we scaled the values of seed production for all
the species with mean 0 and sd of 1. To test the effect of
heterospecific pollen, we substracted to the seed set of hand cross
pollination the seed set of heterospecific pollen treatments. In order
to see correlations between heterospecific pollen effect and traits we
performed Mantel test between the matrix of heterospecific pollen effect
and the distance matrix of each trait (euclidean distances). Moreover,
Mantel test was also conducted between heterospecific pollen effect and
the square root of the matrix of phylogenetic distance due to
improvement in the statistical power (Letten \& Cornwell 2014).
\href{}{all is here, but I would break it by questions and give a bit
more detail, to avoid overwhelm the reader} We explored also the
relations between traits and heterospecific pollen effect through
generalized mixed models where the response variable was heterospecific
pollen effect, the independent variable the different traits and the
random effects the different treatments per species {[}Here I think you
should think if this controls for the non independency of donors and
recipients. I think not. Maybe look onto matrix regresions?). Moreover,
pairwise evolutive distances were calculated with MEGA7 for two kinds of
markers: 1) Internal transcribed spacer (ITS) and 2)
ribulose-bisphosphate carboxylase (RBCL). The sequences of interest were
downloaded from NCBI GenBank and the phylogenetic tree constructed by
maximum likelihood with MEGA7.\href{}{Make a section on how you
contrsucted phylogeny}.

\href{}{I would explain three test. 0) treatment effects with GLM's, 1)
Mantels: relative effects, 2) GLM's or matrix models: Absolute effects
and explain them in three independent paragraphs including a rationale
of why}

>>>>

Pollinator sharing can have negative consequences for plant fitness with
the arrival of foreign pollen. However, the costs of heterospecific
pollen are not yet well understood. We conducted a glasshouse experiment
to understand how phylogenetic relatedness and plant traits mediate the
impacts of heterospecific pollen transfer. We conducted 1800 crosses by
experimentally transferring pollen (50\% and 100\% ratio) with
reciprocal crosses between 10 species belonging to three different
families: Brassicaceae, Solanaceae and Convolvulaceae. Seed set was used
as proxy of plant fitness. We found that for 65\% of the treatments with
50\% mix reduced seed set. Moreover, the reduction in seed set was
dependent on the degree of relatedness and reproductive traits of the
pollen recipient and not the pollen donor. Our results show that certain
traits, particularly compatibility system, are critical in understanding
the costs of heterospecific pollen. \vspace{3mm}\hrule

@@

The study was conducted in a glasshouse at University of New England
(Armidale, Australia) from November 2017 to March 2018. Rooms were
temperature controlled depending on the requirements of the species with
day and night temperature differences. The experimental design had
species from three different families: Solanaceae, Brassicaceae and
Convolvulaceae (\textbf{Table 1}). The species of the study had
different reproductive traits and different degree of relatedness where
the reciprocal crosses between species allowed us to have multiple
different scenarios of both traits and relatedness. Moreover, the
species selected had fast life cycle and low structural flower
complexity in order to perform the pollination treatments and grow the
different species from seeds. For the purpose of the experiment all the
species where considered as pollen recipient and as pollen donor (see
interaction matrix, fig 2). Species were watered once or twice per day
and fertilized weekly (NPK 23: 3.95: 14).

@@

The traits measured for each species were pollen per anther, number of
ovules and stigma, style, ovary width and length. For the stigma, the
stigmatic area was also measured and moreover the stigmas were divided
in wet/dry type. All the morphometrical measurements were performed with
a stereophotomicroscope. Pollen was counted for 20 anthers of each
species with 4 replicates per sample with an hemocytometer. Previously,
anthers were squashed on a known solution with the pippete tip and
homogeneize with a vortex for 30 seconds. Ovule number was counted with
the help of an stereomicroscope and a small grid over a petri dish from
15 randomly selected flowers. Fruits per number of flowers treated were
counted for Solanaceae species and for the rest, the number of seeds
produced per average number of ovules of each species. Levels of self
incompatibility were estimated by dividing the fruit set of hand self
pollination by hand cross pollination Lloyd and Schoen (1992).

@@

Moreover, Mantel test was also conducted between heterospecific pollen
effect and the square root of the matrix of phylogenetic distance due to
improvement in the statistical power (Letten \& Cornwell 2014).
\href{}{all is here, but I would break it by questions and give a bit
more detail, to avoid overwhelm the reader} We explored also the
relations between traits and heterospecific pollen effect through
generalized mixed models where the response variable was heterospecific
pollen effect, the independent variable the different traits and the
random effects the different treatments per species {[}Here I think you
should think if this controls for the non independency of donors and
recipients. I think not. Maybe look onto matrix regresions?). Moreover,
pairwise evolutive distances were calculated with MEGA7 for two kinds of
markers: 1) Internal transcribed spacer (ITS) and 2)
ribulose-bisphosphate carboxylase (RBCL). The sequences of interest were
downloaded from NCBI GenBank and the phylogenetic tree constructed by
maximum likelihood with MEGA7.\href{}{Make a section on how you
contrsucted phylogeny}.

To evaluate heterospecific pollen effect on seed production we performed
linear mixed models. The distinct heterospecific pollination treatments
were compared thorugh relevelling each variable with the cross
pollination treatment which was our control for optimum seed production
for all the species. The different replicates of each treatment were
considered as random effects. The changes of seed production across
species were compared through scaling the seed production with mean 0
and standard deviation of 1 prior to the analysis. All the analysis were
conducted with the statistical language \texttt{R} (R Core Team 2018)

In order to test the relative effect of traits on seed production with
foreign pollen we performed mantel test between the assymetrical matrix
of heterospecific pollen effect (10 by 10 matrix) with the different
distance matrices of traits (euclidean distances). Heterospecific pollen
effect was considered as the subtraction of seed production by hand
cross pollination minus seed production of the different heterospecific
pollen treatments.

The
